<<<<<<< HEAD
---
title: "sound_hab_landscape_metrics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

```{r}
library(SDMTools)
library(sp)
library(raster)
library(rgdal)
library(dplyr)
library(tidyr)
library(ggplot2)
library(rasterVis)

#install.packages("rgeos")
```

## Seagrass metrics
- Exported raster with a pixel size of 3.2 ft (equivalent 1m)
Landscape metrics required meter resolution

```{r}
#SG1 - manually identified habitats
SG1_class = raster('C:/Users/Becca/Documents/git/sound_habitat_2019/data/classified_raster/Classification_Seagrass_1_image2.tif')
#plot(SG1_class)
# calculate metrics
SG_1_patch = PatchStat(SG1_class, cell = 1) 
SG_1_class = ClassStat(SG1_class, cell = 1)
# manually add habitat classification titles from arcmap order 
SG_1_patch = SG_1_patch %>% 
  mutate(habitat = c("mudflat","seagrass", "saltmarsh", "oyster","sand"), site= "SG1")

#SG2 - manually identified habitats
SG2_class = raster("data/classified_raster/Classification_Seagrass_2_image.tif")
#plot(SG2_class)
SG_2_patch = PatchStat(SG2_class, cell = 1)
SG_2_patch = SG_2_patch %>%
  mutate(habitat = c("seagrass", "sand", "channel", "saltmarsh", "oyster"), site= "SG2")

#SG 4 - manually identified habitats
SG4_class = raster("data/classified_raster/Classification_Seagrass_4_image.tif")
#plot(SG4_class)
SG_4_patch = PatchStat(SG4_class, cell = 1)
SG_4_patch = SG_4_patch %>%
  mutate(habitat = c("land", "seagrass", "channel", "sand", "oyster"), site= "SG4")

#SG 6 
SG6_class = raster("data/classified_raster/Classification_Seagrass_6_image.tif")
#plot(SG6_class)
SG_6_patch = PatchStat(SG6_class, cell = 1)
SG_6_patch = SG_6_patch %>%
  mutate(habitat = c("sand","seagrass", "oyster","saltmarsh"), site= "SG6")

#SG 9
SG9_class = raster("data/classified_raster/Classification_Seagrass_9_image.tif")
#plot(SG9_class)
SG_9_patch = PatchStat(SG9_class, cell = 1)
SG_9_patch = SG_9_patch %>%
  mutate(habitat = c("sand", "saltmarsh", "oyster", "seagrass"), site= "SG9")

#SG 10
SG10_class = raster("data/classified_raster/Classification_Seagrass_10_image.tif")
#plot(SG10_class)
SG_10_patch = PatchStat(SG10_class, cell = 1)
SG_10_patch = SG_10_patch %>%
  mutate(habitat = c("sand", "seagrass"), site= "SG10")

#SG 12
SG12_class = raster("data/classified_raster/Classification_Seagrass_12_image.tif")
#plot(SG12_class)
SG_12_patch = PatchStat(SG12_class, cell = 1)
SG_12_patch = SG_12_patch %>%
  mutate(habitat = c("saltmarsh","oyster", "sand","seagrass"), site= "SG12")

#SG 14
SG14_class = raster("data/classified_raster/Classification_Seagrass_14_image.tif")
#plot(SG14_class)
SG_14_patch = PatchStat(SG14_class, cell = 1)
SG_14_patch = SG_14_patch %>%
  mutate(habitat = c("sand","seagrass"), site= "S14")

#SG 15 - manually identified habitats
SG15_class = raster("data/classified_raster/Classification_Seagrass_15_image.tif")
#plot(SG15_class)
SG_15_patch = PatchStat(SG15_class, cell = 1)
SG_15_patch = SG_15_patch %>%
  mutate(habitat = c("oyster", "seagrass", "saltmarsh", "land", "sand"), site= "S15")

#SG 17
SG17_class = raster("data/classified_raster/Classification_Seagrass_17_image.tif")
#plot(SG17_class)
SG_17_patch = PatchStat(SG17_class, cell = 1)
SG_17_patch = SG_17_patch %>%
  mutate(habitat = c("land","saltmarsh","oyster","seagrass","sand"), site= "SG17")

#SG B
SGB_class = raster("data/classified_raster/Classification_Seagrass_B_image.tif")
#plot(SGB_class)
SG_B_patch = PatchStat(SGB_class, cell = 1)
SG_B_patch = SG_B_patch %>%
  mutate(habitat = c("seagrass","saltmarsh","sand"), site= "SB")

#SG C
SGC_class = raster("data/classified_raster/Classification_Seagrass_C_image.tif")
#plot(SGC_class)
SG_C_patch = PatchStat(SGC_class, cell = 1)
SG_C_patch = SG_C_patch %>%
  mutate(habitat = c("seagrass","sand"), site= "SC")

```

## Creek

```{r}

# C2
C2_class = raster("data/classified_raster/Classification_Creek_2_image.tif")
#plot(C2_class)
C_2_patch = PatchStat(C2_class, cell = 1)
C_2_patch = C_2_patch %>%
  mutate(habitat = c("saltmarsh", "seagrass", "oyster", "sand"), site= "C2")

# C3
C3_class = raster("data/classified_raster/Classification_Creek_3_image.tif")
#plot(C3_class)
C_3_patch = PatchStat(C3_class, cell = 1)
C_3_patch = C_3_patch %>%
  mutate(habitat = c("sand", "seagrass", "oyster", "saltmarsh"), site= "C3")

# C4
C4_class = raster("data/classified_raster/Classification_Creek_4_image.tif")
#plot(C4_class)
C_4_patch = PatchStat(C4_class, cell = 1)
C_4_patch = C_4_patch %>%
  mutate(habitat = c("saltmarsh", "oyster", "sand", "channel"), site= "C4")

# C5 - manually identified habitats
C5_class = raster("data/classified_raster/Classification_Creek_5_image.tif")
#plot(C5_class)
C_5_patch = PatchStat(C5_class, cell = 1)
C_5_patch = C_5_patch %>%
  mutate(habitat = c("saltmarsh", "sand", "seagrass", "oyster"), site= "C5")

# C6
C6_class = raster("data/classified_raster/Classification_Creek_6_image.tif")
#plot(C6_class)
C_6_patch = PatchStat(C6_class, cell = 1)
C_6_patch = C_6_patch %>%
  mutate(habitat = c("land", "seagrass", "saltmarsh", "oyster", "sand"), site= "C6")

# C7 - manually identified habitats
C7_class = raster("data/classified_raster/Classification_Creek_7_image.tif")
#plot(C7_class)
C_7_patch = PatchStat(C7_class, cell = 1)
C_7_patch = C_7_patch %>%
  mutate(habitat = c("saltmarsh", "sand", "land", "mudflat", "oyster"), site= "C7")

# C8 - manually identified habitats
C8_class = raster("data/classified_raster/Classification_Creek_8_image.tif")
#plot(C8_class)
C_8_patch = PatchStat(C8_class, cell = 1)
C_8_patch = C_8_patch %>%
  mutate(habitat = c("saltmarsh", "mudflat", "land", "oyster", "sand", "seagrass"), site= "C8")

# CB
CB_class = raster("data/classified_raster/Classification_Creek_B_image.tif")
#plot(CB_class)
C_B_patch = PatchStat(CB_class, cell = 1)
C_B_patch = C_B_patch %>%
  mutate(habitat = c("seagrass", "saltmarsh", "sand", "oyster"), site= "C_B")

```

# Mudflat

```{r}

# M1
M1_class = raster("data/classified_raster/Classification_Mudflat_1_image.tif")
#plot(M1_class)
M_1_patch = PatchStat(M1_class, cell = 1)
M_1_patch = M_1_patch %>%
  mutate(habitat = c("mudflat", "seagrass", "saltmarsh", "oyster", "sand"), site= "MF1")

# M2 - manually identified habitats
M2_class = raster("data/classified_raster/Classification_Mudflat_2_image.tif")
#plot(M2_class)
M_2_patch = PatchStat(M2_class, cell = 1)
M_2_patch = M_2_patch %>%
  mutate(habitat = c("sand", "oyster", "seagrass", "saltmarsh"), site= "MF2")

# M7
M7_class = raster("data/classified_raster/Classification_Mudflat_7_image.tif")
#plot(M7_class)
M_7_patch = PatchStat(M7_class, cell = 1)
M_7_patch = M_7_patch %>%
  mutate(habitat = c("saltmarsh", "land", "oyster", "seagrass", "sand"), site= "MF7")

# MA
MA_class = raster("data/classified_raster/Classification_Mudflat_A_image.tif")
#plot(MA_class)
M_A_patch = PatchStat(MA_class, cell = 1)
M_A_patch = M_A_patch %>%
  mutate(habitat = c("sand", "channel"), site= "MFA")

```

# Organize to add to ecometrics

```{r}

landAll = rbind(SG_1_patch,SG_2_patch,SG_4_patch,SG_6_patch,SG_9_patch,SG_10_patch,
                SG_12_patch,SG_14_patch,SG_15_patch,SG_17_patch,SG_B_patch, SG_C_patch,
                C_2_patch,C_3_patch,C_4_patch,C_5_patch,C_6_patch,C_7_patch,C_8_patch,C_B_patch,
                M_1_patch,M_2_patch,M_7_patch,M_A_patch)


land_long = landAll %>%
  pivot_longer(cols = c(area, core.area, perimeter, perim.area.ratio, shape.index), 
               names_to = c("metric")) %>%
  data.frame()
land_long = land_long[,c(8:11)]

land_wide = land_long %>%
  pivot_wider(id_cols = "site", names_from = c("habitat","metric"), names_sep = "_", values_from = "value", values_fill = 0) %>%
  mutate(total_hab_area = sum(seagrass_area, saltmarsh_area, oyster_area))

write.csv(land_wide, file = "data/clean_data/landscape_metrics.csv")


```


#### Here to end was exploratory - code above was used in analysis


## Landscape metrics package

```{r}
library(landscapemetrics)
library(landscapetools)
library(raster)

# landscapemetrics functions
#SG 14
SG14_class = raster("data/classified_raster/Classification_Seagrass_14_image.tif")

check_landscape(C4_class)
show_landscape(C4_class, discrete = TRUE)
# classes: saltmarsh = 1, seagrass = 2, sand = 5, oyster = 6

# class area: 
lsm_c_ca(C4_class)
# testing nearest neighbor
lsm_p_enn(C4_class)

# area or class (seagrass vs sand)
lsm_c_ca(test_SG12)
test = calculate_lsm(SG14_class, 'patch')
# what is the unit on 
#value? m^2? I think its hectares?
```

# Testing converting to polygons

```{r}
library(sp)
library(raster)
library(rgeos)

SG14_class = raster("data/classified_raster/Classification_Seagrass_14_image.tif")
plot(SG14_class)

#data = SG14_class@data
#data@values
#SG14_class[SG14_class < 2] = NA
pol = rasterToPolygons(SG14_class, dissolve = TRUE)
plot(pol)


# Testing with a raster that has 2 habitats (seagrass/saltmarsh) rather than 1. 
SGB_class = raster("data/classified_raster/Classification_Seagrass_B_image.tif")
plot(SGB_class)

cellStats(SGB_class, max)
cellStats(SGB_class, min)

# This code is clunky, but it sets the sand classification to NA
test = raster(SGB_class)
bs = blockSize(test)
test = writeStart(test, filename = tempfile(), overwrite = TRUE)
for (i in 1:bs$n) {
  v = getValues(SGB_class, row = bs$row[i], nrows = bs$nrows[i])
  v[v == 14] = NA
  writeValues(test, v, bs$row[i])
}
test = writeStop(test)
cellStats(test, max)

plot(test)

SB_pol = rasterToPolygons(test, dissolve = TRUE)
plot(SB_pol)
dim(SB_pol)

str(SB_pol@data)

library(rgeos)
pol_area = gArea(SB_pol, byid = TRUE) 
# if byid = true, gives the area for each cell from the raster, not actual polygons
# if byid = FALSE, gives the area for seagrass and saltmarsh
str(pol_area)


# eventual code for removing small polygons
world.map <- world.map[world.map@data$AREA > 30000, ]
world.map <- subset(world.map, AREA > 30000)

```

# SDM tools to identify patches

```{r}

#SG 14
SG14_class = raster("data/classified_raster/Classification_Seagrass_14_image.tif")
plot(SG14_class)

spgridTest = as(SG14_class, "SpatialGrid")
spgrid = spgridTest@bbox
plot(spgrid)

# extacting the data frame from the raster
class.df = SG14_class@data



# Data frame of class value for each cell? with its coordinate position
# coordinates/extent are both in weird feet units because of originally imagery. Can I change that?
SG14_df = SG14_class %>%
  as("SpatialPixelsDataFrame") %>%
  as.data.frame()

projmap = crs(SG14_class)

# coerce a polygon into dataframe for use in ggplor2
df = spatialpolygons %>%
  fortify()
  
# can subset with filter or subset: subset(object, feat_name = "")

test2 = ConnCompLabel(class.df)
plot(test2)

#SG_14_patch = PatchStat(SG14_class, cell = 1)
#SG_14_patch = SG_14_patch %>%
#  mutate(habitat = c("sand","seagrass"), site= "S14")

proj4string(SG14_class)
```

Notes in projections!!

- Geographic: lat/long, angular units
- Projected: northings/eastings, linear units

Specify projction in R

proj4string(spatialobject)


change coordinate systems: 

identify target coordinate system: proj4string 
spatial objects: spTransform(spatialobject,target CS)
 look up more codes at spatialreference.org

for rasters:
projectraster(raster, target coordinate, and method for interporlating (nearest neighbor))

